
---
title: "Significant Correlations in EU's COVID-19 Data"
author: "Eric Leisten-butterfly team"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the Necessary Libraries

```{r}
library(dplyr)
```

## Data Loading and Pre-processing

```{r}
# Read the dataset 
data <- read.csv("https://covid.ourworldindata.org/data/owid-covid-data.csv",
                 na.strings = "", header=T)

# Filter for European Union
eu_data <- data %>% filter(location == "European Union")
```

## Determining Significant Correlations

```{r}
# variables to check for correlation
variables <- c("total_cases", "new_cases", 
               "total_deaths", "new_deaths", 
               # ... [rest of your variables here]
               "excess_mortality_cumulative_per_million")

res <- NULL
for(i in 1:(length(variables) - 1)) {
  for(j in (i+1):length(variables)) {
    var1 <- variables[i]
    var2 <- variables[j]
    
    # Filter out rows with NA values 
    temp_data <- na.omit(eu_data[, c(var1, var2)])
    
    if(nrow(temp_data) >= 10) {
      cor_res <- cor.test(temp_data[[var1]], temp_data[[var2]])
      res <- rbind(res, c(var1, var2, cor_res$estimate, cor_res$p.value))
    }
  }
}

# Convert to a data frame 
res_df <- as.data.frame(res)
colnames(res_df) <- c("Variable1", "Variable2", "Correlation", "P_Value")

# Filter for significant correlations
significant_correlations <- res_df %>% filter(as.numeric(P_Value) <= 0.05)

print(significant_correlations)
```

## comments
## From this analysis we can see that COVID data contains many significant correlations. For example, we can see that total cases correlate signficantly with total boosters suggesting that boosters may increase the number of COVID cases in the EU. From this I argue that boosters are bad because they may increase an individuals riskyness resulting in overall more COVID transmission (i.e boosters increase case counts). 

## while the above perspective on risk and boosters may be a factor, because we used multiple tests without testing correction, it is possible that our singificant result is a result of chance. For running multiple coparisons on 50 variables, there are over 1000 tests. With this number of tests the chance of a false positive significant result is essentially 1, with ~60 total expected false positives. This is almost a third of all above significant results, or 33% FDR. 